### 2018-05-30 latex/sc18/algorithm.tex
full revised draft ready

\begin{table*}
\centering
\begin{tabular}{|c|cccc|}
\hline
 & \textbf{Computation} & \textbf{Memory - Local MTTKRP} & \textbf{Communication} & \textbf{Memory - Par.~Algorithm}  \\
\hline
Par-NNCP w/ opt.~proc.~grid & $\frac{IR}{P}$ & $\frac{RI^{1/2}}{P^{1/2}}$& $\frac{NRI^{1/N}}{P^{1/N}}$ & $\frac{NRI^{1/N}}{P^{1/N}}$  \\
Par-NNCP w/ gen.~proc.~grid & & & $R\sum_n \frac{I_n}{P_n}$ & $R\sum_n \frac{I_n}{P_n}$  \\
NoDimTree & $\frac{NIR}{P}$ & $\displaystyle \max_n \frac{RI/I_n}{P/P_n}$  &  & \\
\hline
\end{tabular}
\smallskip
\caption{Leading-order per-iteration costs in terms of computation (flops), communication (words moved), and memory (words).  We ignore constants but omit big-Oh notation for clarity.  The first line corresponds to \cref{alg:Par-NNCP-short} with an optimal choice of processor grid and applying the dimension tree optimization locally.  The second line corresponds to the changes in communication and memory for a general processor grid.  The third line corresponds to the changes in computation and memory if the dimension tree optimization is not applied.}
\label{tab:costs}
\end{table*}

>>>>

\begin{table*}
\centering
\begin{tabular}{|c|cccc|}
\hline
 & \textbf{Computation} & \textbf{Memory - Local MTTKRP} & \textbf{Communication} & \textbf{Memory - Par.~Algorithm}  \\
\hline
Par-NNCP w/ opt.~proc.~grid & $\frac{IR}{P}$ & $\frac{RI^{1/2}}{P^{1/2}}$ & $\frac{NRI^{1/N}}{P^{1/N}}$ & $\frac{NRI^{1/N}}{P^{1/N}}$  \\
%\hline
Par-NNCP w/ gen.~proc.~grid & $\frac{IR}{P}$ & $\frac{RI^{1/2}}{P^{1/2}}$ & $\mathbf{R\sum_n \frac{I_n}{P_n}}$ & $\mathbf{R\sum_n \frac{I_n}{P_n}}$  \\
%\hline
Par-NNCP w/o dim.~tree opt. & $\mathbf{\frac{NIR}{P}}$ & $\displaystyle \mathbf{\max_n \frac{RI/I_n}{P/P_n}}$  &  $\frac{NRI^{1/N}}{P^{1/N}}$ & $\frac{NRI^{1/N}}{P^{1/N}}$ \\
\hline
\end{tabular}
\smallskip
\caption{Leading-order per-iteration costs in terms of computation (flops), communication (words moved), and memory (words).  We ignore constants but omit big-Oh notation for clarity.  The first line corresponds to \cref{alg:Par-NNCP-short} with an optimal choice of processor grid and applying the dimension tree optimization locally.  The second line corresponds to the changes in communication and memory for a general processor grid.  The third line corresponds to the changes in computation and memory if the dimension tree optimization is not applied.}
\label{tab:costs}
\end{table*}

### 2018-05-30 latex/sc18/prelims.tex
minor changes
full revised draft ready

While there are multiple optimization methods to compute NNCP, we will focus on a class of methods that use Block Coordinate Descent (BCD), which is also known as the nonlinear Gauss-Seidel method \cite{Bertsekas1999}.
In BCD, the variables are partitioned into blocks, and each variable block is cyclically updated to optimality with all other blocks fixed.
For details on the convergence properties and comparisons of BCD methods for nonnegative matrix and tensor decomposition problems, see \cite{KHP2014}.
We consider BCD methods for NNCP that choose the entire factor matrices as the blocks, which is also often referred to as Alternating Least Squares.
In this case, every subproblem is a linear nonnegative least squares problem. Formally, the following problem 
is solved iteratively for $n = 1 \cdots N$.

%\SplitN{\label{eqn:bcdnncp}}{
$$\HH^{(n)} \leftarrow \argmin_{\HH \geq 0}  \| \Mn{S}{n} \HH^T - {\Mn{M}{n}}^T \|_F^2$$
%}

>>>>

While there are multiple optimization methods to compute NNCP, we will focus on a class of methods that use Block Coordinate Descent (BCD), which is also known as the nonlinear Gauss-Seidel method \cite{Bertsekas1999}.
In BCD, the variables are partitioned into blocks, and each variable block is cyclically updated to optimality with all other blocks fixed.
For details on the convergence properties and comparisons of BCD methods for nonnegative matrix and tensor decomposition problems, see \cite{KHP2014}.
We consider BCD methods for NNCP that choose the entire factor matrices as the blocks, which is also often referred to as Alternating Least Squares.
In this case, every subproblem is a linear nonnegative least squares problem. Formally, the following problem 
is solved iteratively for $n = 1 \cdots N$:

%\SplitN{\label{eqn:bcdnncp}}{
$$\HH^{(n)} \leftarrow \argmin_{\HH \geq 0}  \| \Mn{S}{n} \HH^T - {\Mn{M}{n}}^T \|_F^2.$$
%}
The BCD iteration is guaranteed to converge to a stationary point, but there is no guarantee that it converges to a global minimum.
The number of outer iterations to convergence is problem dependent but typically ranges from the 10s to the 1000s.
